---
title: "hw5_qe"
author: "Kinsey Reed"
date: "3/29/2021"
output: html_document
---

Included in this homework is an example occupancy modeling dataset obtained from camera trap data across
6 mid-Atlantic states:
Bobcat.csv: detection / non-detection of bobcat at each camera. Each row is a camera, and each column is
a replicate survey (1 day).

p covariates.csv: number of people (divided by 100) photographed. Each row is a camera, and each column
is a replicate survey (1 day). This is a detection covariate.

psi covariates.csv: the single covariate represents the proportion of area within 5km recently disturbed
(e.g., logging, housing development). Treat this as a site-level covariate.

1. Load data and place into an unmarkedFrameOccu object
```{r}
library(unmarked)
data <- read.csv('Bobcat.csv')
head(data)
p_covs <- read.csv('p covariates.csv')
site_covs <- read.csv('psi covariates.csv')

#In order to place into object:
#convert detection data to matrix
data_m <- as.matrix(data)
#for both covariates, make sure categorical variables are factors (none here though)
#detection covariates as a list of data.frames
class(p_covs)#already a data.frame
#put into list
det_covs <- list(
  people = p_covs
)
#site-level covariates just need to be a data.frame, which it already is
class(site_covs)

#put all into object
occu_data <- unmarkedFrameOccu(y = data, 
                               siteCovs = site_covs, 
                               obsCovs = det_covs)
```

2. Fit an occupancy model that assumes conditional detection probability is a function of the number of
people photographed at a site, and occupancy probability is a function of disturbance within 5km.

```{r}
#remember order is: ~ detection covariate (+ * etc) ~ site covariate (Etc)
fit <- occu(formula = ~ people ~ Dist_5km, data = occu_data)
summary(fit)
```

3. Interpret the effect of the number of people photographed at each site on conditional detection probability of bobcat. Verity your interpretation in R.

The 'slope coefficient' (or conditional detection probability) associated with the number of people photographed at each site is the log odds ratio (-0.061) associated with a 1-unit change in the number of people at each site. 

The model is on the log odds scale, so interpretation is the same as the other binomial models and is a Bernoulli random variable.



```{r}
#log odds occupancy and detection probability
coef(fit)

#occupancy probability and detection probability (since on logit-scale) 
plogis(coef(fit))



rho_0 <- plogis(coef(fit)[3] + coef(fit)[4] * 0)
rho_1 <- plogis(coef(fit)[3] + coef(fit)[4] * 1)

log((rho_1 / (1 - rho_1)) / (rho_0 / (1 - rho_0))); coef(fit)[4]


```



4. Re-create the Wald test for the effect of disturbance within 5km. What is the null hypothesis? Do you
reject or fail to reject this null hypothesis?

The null hypothesis is that there is no effect of disturbance within 5km on the whether or not a site is occupied. Because the p-value here is well below 0.05, we can reject that null hypothesis, meaning that the probability of occupancy depends on whether there is an effect of disturbance within 5 km on bobcat site occupation.
```{r}
#coefficient divided by std error
summary(fit)
w <- -23.64 / 4.773
w

2 * pnorm(-1 * abs(w))
#same p value: 7.31e-07
```



5. Predict and plot the effect of disturbance within 5km on bobcat occupancy probability. Do this over
the range of observed disturbance within 5km.


```{r}
#make a data frame for occupancy (psi)
new_psi <- data.frame(Dist_5km = seq(from = min(site_covs$Dist_5km),
                                                    to = max(site_covs$Dist_5km), length.out = 100))

predict <- predict(object = fit, newdata = new_psi, type = 'state')
head(predict)

#need link function here for y value?
plot(y = predict$Predicted, x = new_psi$Dist_5km, xlab = 'Proportion Disturbed Area Within 5km',
     ylab = 'Predicted Occupancy Probability', cex.axis = 1.5, cex.lab = 1.5,
     ylim = c(min(predict$lower), max(predict$upper)), type = 'l')
lines(x = new_psi$Dist_5km, y = predict$lower, lty = 2)
lines(x = new_psi$Dist_5km, y = predict$upper, lty = 2)
```

