---
title: "20210214_QE_HW2"
author: "kms"
date: "2/14/2021"
output: html_document
---
This homework is all about verifying output of a fitted linear model. We will use the same beach dataset from last homework. As before, start by fitting a model with OpenBeach as the response, and include an interaction between Year and BeachID.


1. Calculate the residuals. Use the summary() function applied to your numeric vector of residuals to verify that this matches the Residuals: quantiles reported from the summary of your fitted model. Plot a histogram of your residuals. Do they appear reasonably Gaussian? (we will go into detail on formal model checking later in the semester) (5 points)
```{r q1}
beach <- read.csv("beach.csv")
fit <- lm(formula = OpenBeach ~ BeachID * Year, data = beach)
fit
betas <- coef(fit)
betas
#note that the slope coef changed from hw1 from betas[2] to betas[4] (not sure why)

#calculate residuals
residuals1 <- resid(fit)

summary(residuals1)
summary(fit)
#use summary() to compare numeric residual vector to quartiles in fitted model
#they match
#now plot histogram
hist(residuals1)
#Seems pretty lopsided to me, but I suppose it could be considered "reasonably Gaussian".

```

2. Calculate test statistics for your regression coefficients. Verify by comparing to test statistics reported from model output. (5 points)

``` {r q2}
#test statistics for regression coefficients
#est for slope coef for beach A, B, and C
betas
slope_coefA <- betas[4] #Year (negative)
slope_coefB <- betas[5] #BeachIDB:Year
slope_coefC <- betas[6] #BeachIDC:Year

#est for std error
std_errorA <- summary(fit)[['coefficients']]['Year', 'Std. Error']
std_errorA
std_errorB <- summary(fit)[['coefficients']]['BeachIDB:Year', 'Std. Error']
std_errorC <- summary(fit)[['coefficients']]['BeachIDC:Year', 'Std. Error']

#null_hypothesis
null_hyp <- 0


test_statA <- (slope_coefA - null_hyp) / std_errorA
test_statA #-14.443
test_statB <- (slope_coefB - null_hyp) / std_errorB
test_statB #9.478
test_statC <- (slope_coefC - null_hyp) / std_errorC
test_statC #7.992408
#compare to model output stats
summary(fit)
# t values: A = -14.443, B = 9.478, C = 7.992
```

3. Calculate p-values for your regression coefficients. Verify by comparing to p-values reported from model output. What are the associated null hypotheses? Do you reject or fail to reject these null hypotheses? Note: in the lecture slides I reported test statistics are t-distributed with n − 2 degrees of freedom. More generally, these test statistics are t-distributed with n − k degrees of freedom, where k is the number of regression coefficients in your linear model. (5 points)
```{r q3}
#calc p values for regression coefs
#df = n - k (k is # of regression coef from model)
nrow(beach) #62
 #6 coefficients total in original model
#df = 62 - 6, = 56
pt(-1 * abs(test_statA), df = 56)*2 #1.56e-20
pt(-test_statB, df = 56)*2 #3.07e-13
pt(-test_statC, df = 56)*2 #7.96e-11
 
#compare to summary output
summary(fit) 
# BeachA p = < 2.2e-16 (1.56e-20)
# BeachB p = 3.07e-13
# BeachC p = 7.96e-11

#What is the associated null hyp? Reject or fail to reject?
#The associated null hypothesis is that amount of open beach area does not change significantly over the years.
#Here, we reject that null hypothesis. 
```

4. Select a single regression coefficient (your choice) and devise a null hypothesis that is different from the default in lm(). Report the test statistics, your p-value, and whether you reject or fail to reject your null hypothesis. (5 points)
```{r q4}
#Using the slope coefficient for Beach A, I want to test the null hypothesis that B1 = 108 (i.e. the median open beach area) 
#against the alternative that B1 < 108 (that a random value is under 108 ha)
ts_108 <- (betas[4] - 108) / std_errorA
ts_108 # test statistic = -99.38

pt(ts_108, df = 56)
#p val = 5.73e-65

#So here, we reject the null hypothesis. 

```
5. Interpret output of your fitted model. Tell me how beach area does (or does not change) through time at each of the 3 beaches.

```{r q5}
#essentially the same question as q5 in hw1?
summary(fit)
fit
#BeachA
#rate of change (slope) is equal to:
betas[4]
#= -18.36 which means that Beach A looses approximately 18.36 ha of open beach each year.
#BeachB
#rate of change (slope) is equal to:
betas[4] + betas[5]
#= -1.32 meaning that Beach B looses approximately 1.32 ha of open beach each year.
#BeachC
#rate of change (slope) is equal to :
betas[4] + betas[6]
#=-3.99 meaning that Beach C looses approximately 3.99 ha of open beach each year.    

#all have p-values <<< 0.05 so reject the null hypothesis that the beaches do not lose open beach area over time. There is a significant loss of open beach area over time in all 3 beaches, although Beach A is much more severe.
```
