---
title: "20210412HW7_QE"
author: "Kinsey Reed"
date: "4/12/2021"
output: html_document
---

Included in this homework is the same bobcat occupancy modeling dataset from homework 5:
Bobcat.csv: detection / non-detection of bobcat at each camera. Each row is a camera, and each column is
a replicate survey (1 day).
p covariates.csv: number of people (divided by 100) photographed. Each row is a camera, and each column
is a replicate survey (1 day). This is a detection covariate.
psi covariates.csv: the single covariate represents the proportion of area within 5km recently disturbed
(e.g., logging, housing development). Treat this as a site-level covariate.


1. Load data and place into an unmarkedFrameOccu object


```{r}
library(unmarked)
data <- read.csv('Bobcat.csv')
head(data)
p_covs <- read.csv('p covariates.csv')
site_covs <- read.csv('psi covariates.csv')

#In order to place into object:
#convert detection data to matrix
data_m <- as.matrix(data)
#for both covariates, make sure categorical variables are factors (none here though)
#detection covariates as a list of data.frames
class(p_covs)#already a data.frame
#put into list
det_covs <- list(
  people = p_covs
)

#site-level covariates just need to be a data.frame, which it already is
class(site_covs)

#put all into object
occu_data <- unmarkedFrameOccu(y = data, 
                               siteCovs = site_covs, 
                               obsCovs = det_covs)
```

2. Fit the following candidate set of models:


detection intercept-only + occupancy intercept-only
```{r}
fit1 <- occu(formula = ~ 1 ~ 1, data = occu_data)
summary(fit1)
```

detection people + occupancy intercept-only
```{r}
fit2 <- occu(formula = ~ people ~ 1, data = occu_data)
summary(fit2)
```

detection intercept-only  + occupancy disturbance
```{r}
fit3 <- occu(formula = ~ 1 ~ Dist_5km, data = occu_data)
summary(fit3)
```

detection people + occupancy disturbance
```{r}
fit4 <- occu(formula = ~ people ~ Dist_5km, data = occu_data)
summary(fit4)
```

3. Perform model selection with AIC. What is your top model? How do you know? Is there model selection
uncertainty?

The top model is detection: intercept-only and occupancy: disturbance. We know because it is listed first and has a delta AIC = 0. However, the model that includes detection:people and occupancy:disturbance has a delta AIC < 2, so it also offers substantial empirical support for the top model. Because of this, there is model selection uncertainty. 
```{r}
library(AICcmodavg)
cand.list <- list(
  IxI = fit1,
  PxI = fit2,
  IxD = fit3,
  PxD = fit4
)
  
models <- aictab(cand.set = cand.list, second.ord = F)
head(models)
```

4. Average both the effect of people on detection, and disturbance on occupancy, over all models. Report
model-averaged slope coefficients and 95% confidence intervals.
```{r}
#people on detection
avg_P <- modavgShrink(cand.set = cand.list,
             parm = 'people',
             second.ord = F,
             parm.type = 'detect'
             )
print("model averaged slope coefficient for the effect of people on detection")
avg_P$Mod.avg.beta

print("model averaged 95% CI for the effect of people on detection")
avg_P$Lower.CL; avg_P$Upper.CL


#disturbance on occupancy
avg_D <- modavgShrink(cand.set = cand.list,
                      parm = 'Dist_5km',
                      second.ord = F,
                      parm.type = 'psi')

print("model averaged slope coefficient for the effect of disturbance on occupancy")
avg_D$Mod.avg.beta
print('model averaged 95% CI for the effect of disturbance on occupancy')
avg_D$Lower.CL; avg_D$Upper.CL



```

5. Obtain and plot model-averaged predictions of occupancy probability and detection probability. Average
over all models, and make predictions over the observed range of each variable.
```{r}
new.df <- data.frame(
  people = seq(from = min(p_covs, na.rm = TRUE), to = max(p_covs, na.rm = TRUE), length.out = 100),
  Dist_5km = seq(from= min(site_covs$Dist_5km), to = max(site_covs$Dist_5km), length.out = 100)
)

predict_occu <- modavgPred(cand.set = cand.list, 
                    newdata = new.df, 
                    second.ord = F, 
                    parm.type = 'psi')
predict_occu

predict_detect <- modavgPred(cand.set = cand.list, 
                    newdata = new.df, 
                    second.ord = F, 
                    parm.type = 'detect')
predict_detect

plot(y = predict_occu$mod.avg.pred,
     x=new.df$Dist_5km,
     type = 'l',
     ylab = 'Occupancy Probability (Mod.Avgd)',
     xlab='Disturbed Area Proportion (within 5 km)'
     ) 
lines(x = new.df$Dist_5km, y = predict_occu$lower.CL, lty = 2) 
lines(x = new.df$Dist_5km, y = predict_occu$upper.CL, lty = 2)

plot(y = predict_detect$mod.avg.pred,
     x=new.df$people,
     type = 'l',
     ylab = 'Detection Probability (Mod.Avgd)',
     xlab='People'
     )
lines(x = new.df$people, y = predict_detect$lower.CL, lty = 2)
lines(x = new.df$people, y = predict_detect$upper.CL, lty = 2)


#really small confidence intervals?

```

