---
title: "2021030_HW4"
author: "Kinsey Reed"
date: "3/10/2021"
output: html_document
---

y: response variable (counts of the number of events)
x1: continuous predictor variable
x2: categorical predictor variable with 2 levels: "a" and "b"


1. Fit a Poisson regression model that assumes expected count is an interactive function of variables x1
and x2.

```{r}
data <- read.csv('Homework 4 Data.csv')
head(data)

fit <- glm(y ~ x1 * x2, family = poisson, data = data)
summary(fit)
```

2. Interpret the effect of variable x1 on the expected count when x2 is fixed at level "b". Verify your
interpretation in R.

Each additional unit of variable x1, when fixed at level "b", causes a 46.1% increase in the proportional change in expected count.


$ y = \beta_0 + \beta_1x1 + \beta_2x2 + \beta_3x1x2 $
```{r}
B <- coef(fit)
#1-unit change in x1 with x2 fixed at b (1)
c_0 <- B[1] + B[3] 
c_1 <- B[1] + B[2] + B[3] + B[4] 

#log of the ratio of expected counts
log(exp(c_1)/exp(c_0)); B[2] + B[4]

#the proportional change in expected count
exp(c_1) / exp(c_0); exp(B[2] + B[4])

#% increase
(exp(B[2] + B[4])-1)*100
```

3. Interpret the effect of variable x2 on the expected count when x1 is fixed at 1. Verify your interpretation
in R.

There is a 43.3% decrease in the proportional change of expected count between levels a and b,when x1 is fixed at 1.

```{r}
#1-unit change in x2 with x1 = 1
ct_0 <- B[1] + B[2]
ct_1 <- B[1] + B[2] + B[3] + B[4]  

#log of the ratio of expected counts
log(exp(ct_1)/exp(ct_0)); B[3] + B[4] 

#the proportional change in expected count
exp(ct_1) / exp(ct_0); exp(B[3] + B[4])

#% increase
(exp(B[3] + B[4])-1)*100
```


4. Predict the expected count, ± 95% confidence intervals, over the observed range of values of x1, assuming
x2 is fixed at level "b".

```{r}
#data frame of predicted values
nd <- data.frame(
  x1 = seq(min(data$x1), max(data$x1), length.out = 100),
  x2 = factor(x = rep('b', times = 100),
               levels = c('a', 'b'))
)

#predict
prd <- predict.glm(object = fit, newdata = nd, type = 'link', se.fit = T)
low <- exp(prd$fit - qnorm(0.975) * prd$se.fit)
high <- exp(prd$fit + qnorm(0.975) * prd$se.fit)

#plotting
plot(y = exp(prd$fit), x = nd$x1, xlab = 'x1',
     ylab = 'Expected Count', cex.axis = 1.5, cex.lab = 1.5,
     ylim = c(min(low), max(high)), type = 'l')
lines(x = nd$x1, y = low, lty = 2)
lines(x = nd$x1, y = high, lty = 2)

```



5. Predict the expected count, ± 95% confidence intervals, of levels "a" and "b", assuming x1 is fixed at
it’s mean.

```{r}
#df with x1 held at mean

x1_ = c(mean(data$x1), mean(data$x1))
x2_ = factor(c('a', 'b'))

df <- data.frame(
  x1 = x1_,
  x2 = x2_
  )
#predict expected count 
prd2 <- predict.glm(fit, df, type = 'link', se.fit = T)
low2 <- prd2$fit - qnorm(0.975) * prd2$se.fit 
high2 <- prd2$fit + qnorm(0.975) * prd2$se.fit 

plot(y = plogis(prd2$fit), x = c(1, 2), pch = 16,
ylim = c(min(plogis(low2)), max(plogis(high2))),
xaxt = 'n', xlab = '')
axis(side = 1, at = c(1, 2), labels = c('a', 'b'))
lines(x = c(1, 1), y = plogis(c(low2[1], high2[1])))
lines(x = c(2, 2), y = plogis(c(low2[2], high2[2])))





```

