---
title: "20210407_HW6"
author: "Kinsey Reed"
date: "4/7/2021"
output: html_document
---

Included in this homework is an example N-mixture modeling dataset:

count.csv: Counts from each replicate survey. Each row is a site, and each column is a replicate survey.
Column names j1, j2, j3 represent each replicate survey.

obs_covs.csv: A continuous covariate associated with each replicate survey. Each row is a site, and each
column is a replicate survey. Column names j1, j2, j3 represent each replicate survey.

site_covs.csv: Covariates associated with each site. Each row is a site, and each column a covariate.
Column x1 is a continuous covariate, and column x2 is a categorical covariate with 4 levels: "a", "b",
"c", and "d".

1. Load data and place into an unmarkedFramePCount object

```{r}
library(unmarked)
data <- read.csv('count.csv')
head(data)
obs_covs <- read.csv('obs_covs.csv')
site_covs <- read.csv('site_covs.csv')

#In order to place into object:
#convert detection data to matrix
data_m <- as.matrix(data)
#for both covariates, make sure categorical variables are factors 

#site covs as a data.frames
#already a data.frame, but convert x2 to factor
site_covs <- data.frame(
  x1 = site_covs$x1,
  x2 = factor(site_covs$x2)
)
class(site_covs$x2)


#obs/detection covariates as a list of data.frames
#put into list

det_covs <- list(
  reps = data.frame(obs_covs))


#put all into object
nmix_data <- unmarkedFramePCount(y = data_m, 
                               siteCovs = site_covs, 
                               obsCovs = det_covs)
```

2. Fit an N-mixture model that assumes conditional detection probability is a function of the detection
covariate provided, and expected abundance is a additive function of variables x1 and x2.

```{r}
fit <- pcount(formula = ~ reps ~ x1 + x2,
              data = nmix_data, K = 100)
summary(fit)

```

3. Interpret the effect of x1 on the expected count at each site. Verity your interpretation in R.

x1 = site covariate, (i.e. abundance, so on the log scale. link function = exp(); Poisson interpretation)

The 'slope coefficient' associated with x1 is the log proportional change (0.370) associated with a 1-unit change in x1. 
A.k.a. the log of the ratio of expected counts (abundance). Using the link function exp(), every 1-unit increase in x1 results in an increase of 1.45 expected count.


```{r}
betas <- coef(fit)

x1_0 <- exp(betas[1])
x1_1 <- exp(betas[1] + betas[2])

log(x1_1/x1_0)
betas[2]

exp(betas[2])
```

4. Predict and plot the effect of the supplied detection covariate. Do this over the range of this covariate.
```{r}
#data.frame for predicted detection probability #try 200 too

df <- data.frame(
  reps = seq(from = min(obs_covs), to = max(obs_covs), length.out = 100)
  )

prd_p <- predict(object = fit, newdata = df, type = 'det')

#checking
betas

plogis(betas[6] + betas[7]* min(obs_covs))
head(prd_p)

#all good

plot(x = df$reps, y = prd_p$Predicted,
     ylab = 'Detection Probability',
     xlab = 'Detection Covariate',
     type = 'l'
)
lines(x = df$reps, y = prd_p$lower, lty = 2)
lines(x = df$reps, y = prd_p$upper, lty = 2)
     
```

5. Use contrasts to compare expected abundance between all pairwise levels of variable x2. Obtain p-values
associated with each contrast and tell me whether you reject or fail to reject each null hypothesis tested.

For x2 - a and d we reject the null hypothesis because p is less than any reasonable alpha. For x2 - b, p = 0.06 so if the threshold is p = 0.05 then we fail to reject, but personally I think it is still less than a reasonable alpha. For x2 - c, we fail to reject the null hypothesis.
```{r}

#make contrast matrix (x2a is absorbed into intercept)
x <- matrix(
  c(-1, 0, 1, 0, 0,
    -1, 0, 0, 1, 0,
    -1, 0, 0, 0, 1,
    0, 0, -1, 1, 0,
    0, 0, 0, -1, 1),
    nrow = 5, byrow = T
  )

lin_com <- linearComb(obj = fit, coefficients = x,
                      type = 'state')
lin_com

#wald test
w <- coef(lin_com) / SE(lin_com)
w

#p vals
2 * pnorm(-1 * abs(w))

#x2a reject null, x2b do not reject (.06), x2c do not reject (.83), x2d = reject null
```

