---
title: "HW3"
author: "Kinsey Reed"
date: "2/26/2021"
output: html_document
---

data = "Homework 3 Data"
y: response variable (1 for success, 0 for failure)
x1: continuous predictor variable
x2: categorical predictor variable with 2 levels: "a" and "b"


```{r}
data <- read.csv(file = "Homework 3 Data.csv")
head(data)
```


1. Fit a logistic regression model that assumes the probability of success is an additive function of variables
x1 and x2.

```{r}
fit <- glm(y ~ x1 + x2, family = binomial, data = data)
summary(fit)

betas <- coef(fit)
betas
```

$$y_i = \beta_0 + \beta_1x_1 + \beta_2x_2$$

2. Interpret the effect of variable x1 on the log odds of success. Verify your interpretation in R.

The slope coefficient associated with variable x1 is the log odds ratio (-0.168) associated with a 1-unit change in x1. 


```{r}

p_0 <- plogis(betas[1] + betas[2] * 0)
p_1 <- plogis(betas[1] + betas[2] * 1)

log((p_1 / (1 - p_1)) / (p_0 / (1 - p_0))); betas[2]

```



3. Interpret the effect of variable x2 on the log odds of success. Verify your interpretation in R.

The slope coefficient for variable x2 represents the change in the response associated with a 1-unit change in x2. Since x2 is categorical, the coefficient x2b is the difference in the response between the reference (a) and b. In this log regression, this represents the log odds ratio of a point being "b" relative to a point being "a" or non -"b".  

In this case, that log odds ratio is:

```{r}
p_x2b <- plogis(betas[1] + betas[3]*1) #set x1 = 1
p_x2b
p_x2a <- plogis(betas[1] + betas[3]*0)
p_x2a
log((p_x2b / (1 - p_x2b)) / (p_x2a / (1 - p_x2a))); betas[3]

```

4. Duplicate the Wald Test and p-values for variables x1 and x2 performed by the glm() function. Do
you reject or fail to reject your null hypothesis?

For the variable x1, we fail to reject the null hypothesis. For the variable x2, we reject the null hypothesis.

```{r}

#x1

ts1 <- betas[2] / summary(fit)[['coefficients']]['x1', 'Std. Error']

ts1; summary(fit)[['coefficients']]['x1', 'z value']

2 * pnorm(-1 * abs(ts1), mean = 0, sd = 1); summary(fit)[['coefficients']]['x1', 'Pr(>|z|)']

#x2

ts2 <- betas[3] / summary(fit)[['coefficients']]['x2b', 'Std. Error']
ts2; summary(fit)[['coefficients']]['x2b', 'z value']

2 * pnorm(-1 * abs(ts2), mean = 0, sd = 1); summary(fit)[['coefficients']]['x2b', 'Pr(>|z|)']

```

5. Predict and plot the mean probability of success over the range of values of x1.

```{r}
#range of x1
range_x1 <- seq(from = min(data$x1), to = max(data$x1), length.out = 100)

y <- betas[1] + betas[2] * range_x1 + betas[3]

plot(x = range_x1, y = plogis(y), ylab = 'Mean Probability of Success', xlab = 'x1', cex.axis = 1.5, cex.lab = 1.5, type = 'l')

```

